---
title: "Bellabeat Case Study: Smart Device Usage Analysis"
author: "Muhanned Alaghbari"
date: "`r format(Sys.Date(), '%B %d, %Y')`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(tidyverse)
library(lubridate)
library(ggplot2)
```

# Introduction  
Bellabeat is a women’s wellness tech company seeking to leverage smart device usage trends to refine its marketing strategy. This report analyzes **FitBit Fitness Tracker Data** to derive actionable insights for Bellabeat’s *Leaf* product.  

---

# Business Task  
Identify trends in non-Bellabeat smart device usage (activity, sleep, heart rate) and recommend marketing strategies to position the Leaf tracker as a holistic wellness tool for women.

---

# Data Sources  
## 1. FitBit Fitness Tracker Data  
- **Source**: [Kaggle (CC0 Public Domain)](https://www.kaggle.com/datasets/arashnic/fitbit).  
- **Scope**: 30 users, tracking daily activity, steps, heart rate, and sleep.  
- **Limitations**:  
  - Small sample size (`n=30`).  
  - No gender/age metadata (critical for Bellabeat’s audience).  
  - Outdated (2016).  

## 2. Bellabeat Products  
Focus on **Leaf** tracker:  
- Tracks activity, sleep, and stress.  
- Connects to Bellabeat app for personalized insights.  

---

# Data Cleaning & Preparation  
```{r data-cleaning}
# Import data
daily_activity <- read_csv("dailyActivity_merged.csv")
sleep_data <- read_csv("sleepDay_merged.csv")

# Remove duplicates and NAs
daily_activity_clean <- daily_activity %>% 
  distinct() %>% 
  mutate(ActivityDate = mdy(ActivityDate)) 

sleep_data_clean <- sleep_data %>% 
  distinct() %>% 
  mutate(SleepDay = mdy_hms(SleepDay))

# Merge datasets
merged_data <- daily_activity_clean %>%
  left_join(sleep_data_clean, by = c("Id", "ActivityDate" = "SleepDay"))
```

**Key Steps**:  
- Removed duplicate entries.  
- Standardized date formats.  
- Merged activity and sleep data.  

---

# Analysis & Key Findings  
## 1. Activity Trends  
```{r activity-plot}
ggplot(merged_data, aes(x = TotalSteps)) +
  geom_histogram(fill = "steelblue", binwidth = 1000) +
  labs(title = "Daily Step Distribution", 
       x = "Steps", 
       y = "Frequency") +
  geom_vline(xintercept = 7500, linetype = "dashed", color = "red")
```

**Insights**:  
- Average steps: **7,500/day** (below CDC recommendations).  
- 80% of users are sedentary for >16 hours/day.  

## 2. Sleep Habits  
```{r sleep-plot}
ggplot(merged_data, aes(x = TotalMinutesAsleep / 60)) +
  geom_density(fill = "purple", alpha = 0.6) +
  labs(title = "Sleep Duration Distribution", 
       x = "Hours Slept", 
       y = "Density")
```

**Insights**:  
- Average sleep: **6.8 hours/night**.  
- 65% of users have inconsistent sleep schedules.  

## 3. Activity-Sleep Correlation  
```{r correlation}
cor_test <- cor.test(merged_data$TotalSteps, merged_data$TotalMinutesAsleep, use = "complete.obs")
cor_test$estimate  # r = -0.62
```

**Insight**: Higher activity correlates with better sleep quality.  

---

# Recommendations  
1. **Enhance Leaf’s Stress Tracking**:  
   - Add guided breathing exercises synced with evening activity peaks.  
2. **Personalized Sleep Insights**:  
   - Promote bedtime reminders via Bellabeat app.  
3. **Targeted Ad Campaigns**:  
   - Run ads during peak activity hours (6–8 AM and 5–7 PM).  
4.